A community-maintained standard library of population genetic models

### 2020-04-27 popSimManu.tex
added text about pop-size scaling

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past two decades. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{Boyko:2008cr,kim2017inference,tataru2017inference,Fortier703918,Huang2019,Vecchyo770966},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247,Barroso2019},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While these methods have undoubtedly increased our understanding of
genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of these inferences.
% or their robustness to deviations from the assumptions.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming ever more important.

@@

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run standardized simulations
across a range of organisms. \Stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}.
The tree sequence format could also be converted
to other formats (e.g., VCF) by the user if desired.
Currently, \stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine.
We have implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim}
as an alternative backend,
to allow simulation of processes that cannot be modeled under the coalescent. We validated our
implementation of the \texttt{SLiM} engine by comparing estimates of several population
genetic summary statistics for simulations generated by both engines 
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}).

@@

One of our goals is to engage research communities studying other taxa,
so as to expand the resource to many more species.
Although we have included demographic models and recombination maps,
there are many biological processes that we do not model.
Some of these additions that we are enthusiastic to add are:
selection (including distributions of fitness effects, maps of functional elements, and single sweeps),
gene conversion, mutation models (rate heterogeneity),
more realistic demography (overlapping generations, separate sexes, mortality/fecundity schedules),
geographic population structure,
and downstream aspects of data quality (genotyping and mapping error).
These plans would not be possible without recent important work on SLiM \citep{haller2019slim},
as well as current work on \texttt{msprime}.
Some other important processes are more challenging to model with current simulation software,
such as
structural variation,
changing recombination maps over time,
transposable elements,
and context-dependent mutation.

>>>>

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past two decades. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{Boyko:2008cr,kim2017inference,tataru2017inference,Fortier703918,Huang2019,Vecchyo770966},
recombination rates \citep{chan2012genome,lin2013fast,adrion2020predicting,Barroso2019},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While these methods have undoubtedly increased our understanding of
genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of these inferences.
% or their robustness to deviations from the assumptions.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming ever more important.

@@

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run standardized simulations
across a range of organisms. \Stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}.
The tree sequence format could also be converted
to other formats (e.g., VCF) by the user if desired.
Currently, \stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine.

Coalescent simulations, while highly efficient, are limited in their ability to model
complex selection scenarios, such as recurrent sweeps and background selection. For these reasons
we have also implemented the forward-time simulator, \texttt{SLiM} \citep{haller2019tree,haller2019slim}, 
as an alternative backend engine to \texttt{stdpopsim}, allowing for the simulation of processes that cannot be modeled under 
the coalescent. However, as forward-time simulators explicitly model all individuals in a population, simulating large
population sizes can be highly demanding of computational resources. One solution to this obstacle is to 
rescale the population size, \textit{N}, such that population-dependent parameters (e.g. \textit{N$\mu$}, \textit{Nr},
and \textit{Ns}) remain unchanged \citep[see the \texttt{SLiM} manual and][]{uricchio2014robust}.

We validated our implementation of the \texttt{SLiM} engine by comparing 
estimates of several population genetic summary statistics for neutral simulations generated by both \texttt{SLiM}
and \texttt{msprime} under different factors of population-size rescaling 
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}). We show that rescaling by up to a factor
of $50$ ($Q=50$ in \texttt{SLiM}) produces patterns of both diversity and linkage disequilibrium that are 
qualitatively similar to those observed under the coalescent (i.e. \texttt{msprime}), when both are simulated under a recombination map 
(Figure \ref{fig:slim_val_map}). It is important to note that rescaling \textit{N} is not, in actuality, the equivalent to simulating 
all individuals in the population, and that differences between the two methods may be subtle and difficult to interpret.
To emphasize this point, we show deviations between patterns of diversity generated by \texttt{SLiM} and those generated by 
\texttt{msprime} engines at $Q=50$, when simulating under a uniform recombination rate (Figure \ref{fig:slim_val_nomap}). 
The causes of these differences remain unclear. 
While an in-depth investigation of the effects of population-size rescaling under selection 
scenarios is outside the scope of this report, we are enthusiastic about the potential for \texttt{stdpopsim} and the 
\texttt{SLiM} engine to facilitate addressing such pertinent questions in the near future.

@@

One of our goals is to engage research communities studying other taxa,
so as to expand the resource to many more species.
Although we have included demographic models and recombination maps,
there are many biological processes that we do not model.
Some of these additions that we are enthusiastic to add are:
selection (including distributions of fitness effects, maps of functional elements, and single sweeps),
gene conversion, mutation models (rate heterogeneity),
more realistic demography (overlapping generations, separate sexes, mortality/fecundity schedules),
geographic population structure,
and downstream aspects of data quality (genotyping and mapping error).
Moreover, an in-depth investigation into the effects of population-size rescaling under many of the
above scenarios is warented, given our preliminary findings using neutral simulations 
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}).
These plans would not be possible without recent important work on SLiM \citep{haller2019slim},
as well as current work on \texttt{msprime}.
Some other important processes are more challenging to model with current simulation software,
such as
structural variation,
changing recombination maps over time,
transposable elements,
and context-dependent mutation.
